Change in Mobility and Site Occupation during the Late Pleistocene in Korea

### 2019-11-21 analysis/paper/paper.Rmd
Done with the putting citations

Radiocarbon dates provide a proxy for change in human population levels. Summed probability distributions (SPD) of radiocarbon dates are often used to understand temporal trends in ancient human populations [Rick 1987; Bamforth and Grund 2012; Shennan et al. 2013; Timpson et al. 2014; Contreras and Meadows 2014]. Without there must be unavoidable limitations for the validity of SPD such as the lack of excavated sites in certain period, this method is still helpful to be used as a proxy for population fluctuation. Timpson et al. demonstrate its usefulness by testing data sets in 12 European regions and assert that this SPD is extremely robust to deal with small sample size (Timpson et al. 2014). We use 171 radiocarbon dates from 22 sites to search for population trends during the Late Pleistocene in Korea based on the premise that population change or pressure could be another catalyst to improve or innovate technology.

In terms of the composition of toolkits, we count kind of tools as well as the amount of debitabitage to search for the change in the toolkits around the transition period. The tools represent paleo hunter-gatherers’ adaptive strategies. We analyze the composition of raw materials in the assemblages to see the relationships between the choice of specific raw materials and technological innovation as well as raw material availability and its consuming pattern based on an assumption that stone raw materials were carefully transported and consumed to optimize mobility and technological strategies, depth of planning in landscape use and risk minimization (Bratingham 2003).

@@

Raw materials
During the transition period, the use of shale and rhyolite greatly increased. These two raw materials are suitable to make elongate blades and stemmed points. It is considered that he hunter-gatherer started to use raw materials selectively depending on the type of tool, especially the newly produced stemmed points. Raw material availability is a crucial factor in determining the trip and direction for hunter-gatherers (Davies et al. 2018). Therefore the adoption of new raw material along with the emergence of new type of tools demonstrates a significant innovation in not just technology but also cognitive system.

>>>>

Radiocarbon dates provide a proxy for change in human population levels. Summed probability distributions (SPD) of radiocarbon dates are often used to understand temporal trends in ancient human populations [@rick1987dates; @bamforth2012radiocarbon; @shennan2013regional; @timpson2014reconstructing; @contreras2014summed]. Without there must be unavoidable limitations for the validity of SPD such as the lack of excavated sites in certain period, this method is still helpful to be used as a proxy for population fluctuation. Timpson et al. demonstrate its usefulness by testing data sets in 12 European regions and assert that this SPD is extremely robust to deal with small sample size [@timpson2014reconstructing]. We use 171 radiocarbon dates from 22 sites to search for population trends during the Late Pleistocene in Korea based on the premise that population change or pressure could be another catalyst to improve or innovate technology.

In terms of the composition of toolkits, we count kind of tools as well as the amount of debitabitage to search for the change in the toolkits around the transition period. The tools represent paleo hunter-gatherers’ adaptive strategies. We analyze the composition of raw materials in the assemblages to see the relationships between the choice of specific raw materials and technological innovation as well as raw material availability and its consuming pattern based on an assumption that stone raw materials were carefully transported and consumed to optimize mobility and technological strategies, depth of planning in landscape use and risk minimization [@brantingham2003neutral].

@@

Raw materials
During the transition period, the use of shale and rhyolite greatly increased. These two raw materials are suitable to make elongate blades and stemmed points. It is considered that he hunter-gatherer started to use raw materials selectively depending on the type of tool, especially the newly produced stemmed points. Raw material availability is a crucial factor in determining the trip and direction for hunter-gatherers [@davies2018modeling]. Therefore the adoption of new raw material along with the emergence of new type of tools demonstrates a significant innovation in not just technology but also cognitive system.
